---
title: "Palmer Penguins Notebook"
author: Drew Croteau
output: html_notebook
---

<br> 

#### Purpose
A simple *progression-based* analysis of penguin data.

***


## Creating the R environment:
Here we'll install and load the necessary R packages.

<br>

#### Package installation:
**Note**: If receiving **error(s)** then the packages are most likely *already* installed.
```{r install packages as needed}
install.packages('tidyverse')
install.packages('palmerpenguins')
install.packages('ggplot2')
```

<br>

#### Load packages:
```{r}
library('tidyverse')
library('palmerpenguins')
library('ggplot2')
```

<br>

#### Load and preview the dataset:
**Note**: The 'penguins' dataset is pulled from the 'palmerpenguins' package.
```{r}
data(penguins)
head(penguins)
```

<br>

***

## Cleaning the data:
In the preview of the 'penguins' dataset we can see 'NA' values which require cleaning before manipulating the data and creating visuals.

<br>

#### Creating a new table:
First we'll start by creating a new table and filter the data from 'penguins' to only pull values that are **NOT** 'NA'.
```{r}
clean_penguins = filter(penguins, penguins$body_mass_g != 'NA')
```

<br>

#### View the new table:
Let's take a peek at the new filtered data.
```{r}
head(clean_penguins)
```

<br>

***

## Data manipulation
Now that we have a clean dataset, we can manipulate the data to make visualizations easier to create.

<br>

#### Body mass summary
Here we'll create a subset summary table of our clean dataset with a focus on body mass by species.
```{r average body mass by species}
body_mass_summary = clean_penguins %>% 
  group_by(species) %>%
  summarize(avg_body_mass = mean(body_mass_g))
```

<br>

#### View the summary
Let's view the new summary table and make sure everything is in order.
```{r}
print(body_mass_summary)
```

<br> 

***

## Data visualization
Even though the summary paints a fairly clear picture already, let's make a visualization to show the differences anyway.

<br>

#### Average body mass by species visualization
Utilizing the 'ggplot2' package, we can create an easy-to-understand visual representing the body mass by species summary. In this instance, we'll be using a simple bar chart.
```{r}
ggplot(data = body_mass_summary) %>% 
  + geom_col(mapping = aes(x = species, y = avg_body_mass, fill = species)) %>% 
  ## The rest of this code is for readability and aesthetic purposes.
  + guides(fill = 'none') %>% 
  + labs(title = 'Average Body Mass', x = 'Penguin Species', y = 'Mass (Grams)', caption = 
           'Data by Dr. Kristen Gorman 2007-2009')
```

